Data Flow Analysis\\in the Presence of Correlated Calls

### 2014-12-09 content/intro.tex
related work

Data-flow analysis computes an approximation of how values may flow  
through a program, and has applications in compiler optimization, 
programming tools, and computer security, and many other areas.
Data-flow analyses operate on \textit{control-flow graphs} (CFGs)%
\mtodo{If we introduce this abbreviation, should we replace the uses of
``control-flow graph'' with ``CFG'' in the paper?}
that model 
the order in which the instructions of a program are executed. This is
typically done by associating \textit{flow functions} that represent how 
data is propagated with the edges of the control-flow graph.  A 
\textit{confluence operator} specifies how the data facts that have been 
computed along different paths should be merged when the paths join.

Since a control-flow graph is an over-approximation of the possible flows 
of control in concrete executions of a program, it may contain \textit{infeasible} 
paths that cannot occur at runtime. The precision of a data-flow analysis algorithm
depends critically on its ability to detect and disregard such infeasible paths.
The popular \textit{Interprocedural Finite Distributive Subset} (IFDS) algorithm 
by  Reps, Horwitz, and Sagiv \cite{reps1995precise} is a general data-flow analysis
algorithm for computing solutions to finite distributive data-flow problems such 
as reaching definitions, available expressions, and taint analysis. A distinguishing 
characteristic of IFDS is that it avoids infeasible interprocedural paths in which calls 
and returns to/from functions are not properly matched. Sagiv, Reps, and Horwitz
also presented the \textit{Interprocedural Distributive Environment} (IDE) algorithm
\cite{sagiv1996precise} that similarly only considers properly matched call/return
edges, but that supports a broader range of dataflow problems by expanding the 
domain of flow functions to \textit{environments} that go beyond the data-flow facts 
considered by IFDS.

>>>>

Data-flow analysis computes an approximation of how values may flow  
through a program, and has applications in compiler optimization, 
programming tools, and computer security, and many other areas.
Data-flow analyses operate on \textit{control-flow graphs} (CFGs)
\mtodo{If we introduce this abbreviation, should we replace the uses of
``control-flow graph'' with ``CFG'' in the paper?}
that model 
the order in which the instructions of a program are executed. This is
typically done by associating \textit{flow functions} that represent how 
data is propagated with the edges of the control-flow graph.  A 
\textit{confluence operator} specifies how the data facts that have been 
computed along different paths should be merged when the paths join.

Since a control-flow graph is an over-approximation of the possible flows 
of control in concrete executions of a program, it may contain \textit{infeasible} 
paths that cannot occur at runtime. The precision of a data-flow analysis algorithm
depends critically on its ability to detect and disregard such infeasible paths.
The popular \textit{Interprocedural Finite Distributive Subset} (IFDS) algorithm 
by  Reps, Horwitz, and Sagiv \cite{reps1995precise} is a general data-flow analysis
algorithm for computing solutions to finite distributive data-flow problems such 
as reaching definitions, available expressions, and taint analysis.
\mtodo{Should we include any references here for people who don't know what these analyses refer to?}
\mtodo{We should define somewhere what taint analysis is, since it's currently
commented out.}
A distinguishing 
characteristic of IFDS is that it avoids infeasible interprocedural paths in which calls 
and returns to/from functions are not properly matched. Sagiv, Reps, and Horwitz
also presented the \textit{Interprocedural Distributive Environment} (IDE) algorithm
\cite{sagiv1996precise} that similarly only considers properly matched call/return
edges, but that supports a broader range of dataflow problems by expanding the 
domain of flow functions to \textit{environments} that go beyond the data-flow facts 
considered by IFDS.

### 2014-12-09 content/related.tex
related work

\section{Related Work}
IFDS is a version of the functional approach to data-flow analysis developed by M.\,Sharir and A.\,Pnueli~\cite{pnueli1981two}. Their algorithm is based on computing \textit{summary functions} that return the data-flow value at the end of a procedure, given the data-flow value at the start of the procedure. IFDS problems form a more restricted set of data-flow problems: unlike in the functional approach, IFDS flow functions have to be distributive, and the set of data-flow facts $D$ has to be finite. However, the IFDS algorithm is more general than Sharir's and Pnueli's algorithm in that it can handle programs containing local variables and parameters in recursive methods.

IFDS has been used to encode a variety of data-flow problems. More complex examples of applications include typestate analysis (determining which operations can be performed on an object at a given program point)~\cite{naeem2008typestate} or shape analysis (detecting errors and validating properties of programs at compile time)~\cite{DBLP:conf/birthday/KreikerRRSWY13}.

IFDS is implemented for two popular static-analysis frameworks, the T.J. Watson Libraries for Analysis (WALA)~\cite{fink2012wala} and Soot~\cite{vallee1999soot}.

WALA is a framework for static analysis on Java bytecode developed by the IBM T.J.~Watson Research Center.
In the implementation of our work, we use WALA to build and traverse the supergraph (a special kind of control-flow graph) of a Java program\footnote{However, we do not use WALA's IFDS implementation, as explained in~Section~\ref{chapter:eval}.}.

@@

Whereas one advantage of Soot's IFDS implementation (and other static analysis tools) is ease of use and extensibility, WALA's primary focus is efficiency. For example, WALA uses bit-vectors to represent some of the analysis data types, like local variables and parameters. Another difference is that WALA's intermediate representation of a program uses static single assignment (SSA) form~\cite{cytron1991efficiently}. SSA form is a representation of the program in which each variable has only one definition (assignment). SSA can make dataflow analysis simpler and more efficient~\cite{appel1998ssa}.

Work on improving the IFDS algorithm includes Practical Extensions by~N.\,Naeem and O.\,Lhot\'ak~\cite{naeem2010pei}. Their paper presents four extensions to the IFDS algorithm. Two of the extensions improve the efficiency of the IFDS analysis for certain classes of IFDS problems. Another extension widens the class of problems applicable for the IFDS analysis. However, those extensions do not affect the precision of IFDS problems. Our analysis, in contrast, does not improve the efficiency or generality of IFDS, but it allows us to solve IFDS problems more precisely.

The fourth extension is targeted towards programs that are represented in SSA form. Executing the IFDS analysis on such programs results in loss of precision in the presence of control-flow constructs (e.g. conditionals and loops), compared to programs in non-SSA form.
The extension makes the IFDS analysis on programs in SSA form as precise as on programs that are not represented in SSA form. In contrast, the correlated-calls analysis is applicable to programs in both SSA and non-SSA forms. Even if applied to a program in SSA form, our analysis and the extension improve the precision of IFDS in unrelated situations: the first analysis handles correlated calls, and the latter handles control-flow constructs. Thus, an IFDS analysis could benefit from both precision improvements independently.

@@

The correlated-calls analysis improves the precision of a data-flow analysis by eliminating a special type of infeasible paths. This is similar to the idea of context-sensitive analysis: just as a context-sensitive analysis eliminates infeasible paths from the end of a procedure to the call sites that do not match the given procedure call, the correlated-calls analysis eliminates infeasible paths caused by correlated method calls.

>>>>

\section{Related Work}
IFDS is a version of the functional approach to data-flow analysis developed by M.\,Sharir and A.\,Pnueli~\cite{pnueli1981two}.
\commentout{
Their algorithm is based on computing \textit{summary functions} that return the data-flow value at the end of a procedure, given the data-flow value at the start of the procedure. IFDS problems form a more restricted set of data-flow problems: unlike in the functional approach, IFDS flow functions have to be distributive, and the set of data-flow facts $D$ has to be finite. However, the IFDS algorithm is more general than Sharir's and Pnueli's algorithm in that it can handle programs containing local variables and parameters in recursive methods.
}
IFDS has been used to encode a variety of data-flow problems, for~example, typestate analysis%
\commentout{(determining which operations can be performed on an object at a given program point)}~\cite{naeem2008typestate,DBLP:conf/pldi/ZhangMGNY14} or shape analysis%
\commentout{(detecting errors and validating properties of programs at compile time)}
~\cite{DBLP:conf/birthday/KreikerRRSWY13}. IFDS has been broadly used~\cite{DBLP:conf/pldi/ArztRFBBKTOM14,tripp2009taj} and extended~\cite{DBLP:conf/sigsoft/LerchHBM14} to solve taint-analysis problems.

IFDS is implemented for two popular static-analysis frameworks for Java bytecode, the T.J.~Watson Libraries for~Analysis (WALA)~\cite{fink2012wala} and Heros~\cite{bodden2012inter}.

\commentout{
WALA is a framework for static analysis on Java bytecode developed by the IBM T.J.~Watson Research Center.
In the implementation of our work, we use WALA to build and traverse the supergraph (a special kind of control-flow graph) of a Java program\footnote{However, we do not use WALA's IFDS implementation, as explained in~Section~\ref{chapter:eval}.}.

@@

Whereas one advantage of Soot's IFDS implementation (and other static analysis tools) is ease of use and extensibility, WALA's primary focus is efficiency. For example, WALA uses bit-vectors to represent some of the analysis data types, like local variables and parameters. Another difference is that WALA's intermediate representation of a program uses static single assignment (SSA) form~\cite{cytron1991efficiently}. SSA form is a representation of the program in which each variable has only one definition (assignment). SSA can make dataflow analysis simpler and more efficient~\cite{appel1998ssa}.
}

Work on improving the IFDS algorithm includes the Practical Extensions to the IFDS algorithm~\cite{naeem2010pei}.
\commentout{The paper presents four extensions to the IFDS algorithm.}
Two of the four extensions improve the efficiency of the IFDS analysis for certain classes of IFDS problems. Another extension widens the class of problems applicable for the IFDS analysis. 
\commentout{However, those extensions do not affect the precision of IFDS problems.}
Our analysis, in contrast, does not improve the efficiency or generality of IFDS, but it allows us to solve IFDS problems more precisely.
The fourth extension is targeted towards programs that are represented in SSA form. Executing the IFDS analysis on such programs results in loss of precision in the presence of control-flow constructs (e.g. conditionals and loops), compared to programs in non-SSA form.
The extension makes the IFDS analysis on programs in SSA form as precise as on programs that are not represented in SSA form. In contrast, the correlated-calls analysis is applicable to programs in both SSA and non-SSA forms. Even if applied to a program in SSA form, our analysis and the extension improve the precision of IFDS in unrelated situations: the first analysis handles correlated calls, and the latter handles control-flow constructs. Thus, an IFDS analysis could benefit from both precision improvements independently.

@@

\commentout{
The correlated-calls analysis improves the precision of a data-flow analysis by eliminating a special type of infeasible paths. This is similar to the idea of context-sensitive analysis: just as a context-sensitive analysis eliminates infeasible paths from the end of a procedure to the call sites that do not match the given procedure call, the correlated-calls analysis eliminates infeasible paths caused by correlated method calls.
}
